---
title: "Lipids isolated effects on coefficients"
author: "Luís Romera"
format: html
editor: visual
---

```{r}
getwd()
```

In this document, I'm going to calculate the mean of the coefficients of each lipid used in the nanostructure formulations.

## Lipids:

- **Bacuri**

- **Açaí**

- **Pracaxi**

- **Cacau**

- **Tucumã**

- **Cupuaçu**

- **Pequi**

- **Manga**

- **Andiroba**

- **Murumuru**

- **Buriti**

- **Jambu**

- **Urucum**

- **Ucuuba**

I'm going to apply multiple linear regression to determine how each lipid dictates the **Scheffé's coefficients**.

```{r}
coefData <- read.csv("Tabela de coeficientes NanoIA.csv")
```

Let's fit the linear models

## B1

```{r}
B1 <- lm(
  b1 ~
    Bacuri +
    Açaí +
    Pracaxi +
    Cacau +
    Tucumã +
    Cupuaçu +
    Pequi +
    Manga +
    Andiroba +
    Murumuru +
    Buriti +
    Jambu +
    Urucum +
    Ucuuba,
  data = coefData
)

B1

summary(B1)

anova(B1)
```

## B2

```{r}

lm(
  b2 ~
    Bacuri +
    Açaí +
    Pracaxi +
    Cacau +
    Tucumã +
    Cupuaçu +
    Pequi +
    Manga +
    Andiroba +
    Murumuru +
    Buriti +
    Jambu +
    Urucum +
    Ucuuba,
  data = coefData
)
```

## B3

```{r}
lm(
  b3 ~
    Bacuri +
    Açaí +
    Pracaxi +
    Cacau +
    Tucumã +
    Cupuaçu +
    Pequi +
    Manga +
    Andiroba +
    Murumuru +
    Buriti +
    Jambu +
    Urucum +
    Ucuuba,
  data = coefData
)
```

## B12

```{r}
B12 <- lm(
  b12 ~
    Bacuri +
    Açaí +
    Pracaxi +
    Cacau +
    Tucumã +
    Cupuaçu +
    Pequi +
    Manga +
    Andiroba +
    Murumuru +
    Buriti +
    Jambu +
    Urucum +
    Ucuuba,
  data = coefData
)

B12

summary(B12)
```

## B13

```{r}

lm(
  b13~
    Bacuri +
    Açaí +
    Pracaxi +
    Cacau +
    Tucumã +
    Cupuaçu +
    Pequi +
    Manga +
    Andiroba +
    Murumuru +
    Buriti +
    Jambu +
    Urucum +
    Ucuuba,
  data = coefData
)
```

## B23

```{r}

lm(
  b23~
    Bacuri +
    Açaí +
    Pracaxi +
    Cacau +
    Tucumã +
    Cupuaçu +
    Pequi +
    Manga +
    Andiroba +
    Murumuru +
    Buriti +
    Jambu +
    Urucum +
    Ucuuba,
  data = coefData
)
```

Now, I'm going to apply PCA to the lipids alone.

```{r}

getwd()

coefSing <- read.csv("coeficientesSingular.csv")

coefSing <- coefSing[rowSums(is.na(coefSing)) < ncol(coefSing), ]

coefSing <- coefSing[, colSums(is.na(coefSing)) < nrow(coefSing)]

coefSing <- coefSing[!is.na(coefSing$Lipid), ]

coefSing <- coefSing[coefSing$Lipid != "", ]

nrow(coefSing)

str(coefSing)

head(coefSing)

pcaData <- coefSing[, c(
"b1",
"b2",
"b3",
"b12",
"b13",
"b23"
)]

rownames(pcaData) <- coefSing$Lipid

View(coefSing)

pca <- prcomp(
  pcaData,
  retx = TRUE,
  center = TRUE,
  scale. = TRUE
)
pca

summary(pca)

library(factoextra)

pcaEigLip <- fviz_eig(pca)
pcaEigLip

ggsave(
  "pcaEigLip.png",
  plot = pcaEigLip,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 300
)

pcaBiplot <- fviz_pca_biplot(
  pca,
  repel = TRUE,
  pointsize = 2
)
pcaBiplot

ggsave(
  "pcaBiplot.png",
  plot = pcaBiplot,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 300
)

pcaVar <- fviz_pca_var(
  pca,
  col.var = "contrib",
  repel = TRUE,
  gradient.cols = c("blue", "orange", "red")
)
pcaVar

ggsave(
  "pcaVar.png",
  plot = pcaVar,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 300
)

round(pca$rotation, 3)
```

## Clustering

```{r}

distMat <- dist(scale(pcaData))

hc <- hclust(
  distMat,
  method = "ward.D2"
)

hc

cluster <- cutree(
  hc,
  k = 2
)
cluster

pcaHcLipid <- fviz_pca_ind(
  pca,
  geom = "point",
  habillage = factor(cluster),
  addEllipses = TRUE,
  ellipse.level = 0.95,
  pointsize = 4,
  repel = TRUE
) +
  theme_classic()

pcaHcLipid

ggsave(
  "pcaHcLipid.png",
  plot = pcaHcLipid,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 300
)

pcaComp <- fviz_pca_biplot(
  pca,
  habillage = factor(cluster),
  addEllipses = TRUE,
  ellipse.level = 0.95,
  pointsize = 4,
  repel = TRUE
) 

#pcaComp

pcaCluster <- fviz_cluster(
  list(
    data = scale(pcaData),
    cluster = cluster
  )
)

pcaCluster

ggsave(
  "pcaCluster.png",
  plot = pcaCluster,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 300
)
```

## Dendrogram

```{r}

#library(factoextra)
library(dendextend)

dend <- as.dendrogram(hc)

dend <- color_branches(
  dend,
  k = 3
)

plot(dend)

class(dend)

#png(
 # "dend.png",
 # width = 18,
#  height = 14,
 # units = "cm",
#  res = 300
#)

plot(
  dend,
  main = "Cluster Dendrogram",
  xlab = "",
  sub = "",
  cex = 1.2
)

dev.list()

dev.off()
```

## Heatmap

```{r}

library(pheatmap)

pcaDataScale <- scale(pcaData)

heatMap <- pheatmap(
  pcaDataScale,
  clustering_method = "ward.D2",
  border_color = NA,
  fontsize_row = 12,
  fontsize_col = 12,
  main = "Lipid clustering by Scheffé's Coefficients"
)

class(heatMap)

heatMap
```
