---
title: "Lipids Analysis"
author: "Luís Romera"
format: pdf
editor: visual
---

In this document, I'm going to analyse the relationship between the lipids used for formulation of nanoparticles.

## PCA of the lipid pairs

First, I'm going to make a principal component analysis of the models coefficients and see what parameter each lipid pair approximates.

```{r}
#| dev: png


library(magick)
coefData <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/Topicos_em_Nanotecnologia/Tabela de coeficientes NanoIA.csv")

head(coefData)

coefDataPca <- coefData[, c("b1", "b2", "b3", "b12", "b13", "b23")]

rownames(coefDataPca) <- coefData$Combination

coefPca <- prcomp(
  x = coefDataPca,
  retx = TRUE,
  center = TRUE,
  scale. = TRUE
)

coefPca

summary(coefPca)
```

Now, lets try to do some plots,

starting with the scree plot of the explained variance per dimension.

```{r}
#| dev: png

#| warning: False
#| message: False

library(factoextra)

screePlot <- fviz_eig(coefPca,
                     # choice = "eigenvalue",
                      barcolor = "black",
                      addlabels = TRUE,
                      ggtheme = theme_minimal(),
                    #  parallel = TRUE
                    ) #+ theme_void()

screePlot

ggsave(
  "pcaCoefVar.png",
  plot = screePlot,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 600
)
```

First, I'll plot the parameter's vectors

```{r}
#| dev: png

PcaVar <- fviz_pca_var(
  X = coefPca,
  repel = TRUE,
  arrowsize = 0.8,
  col.var = "contrib",
  gradient.cols = c("#40B5AD", "#FA5F55")
) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "PCA of Scheffé model's coefficients"
  )


PcaVar

ggsave(
  "pcaVar.png",
  plot = PcaVar,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 600
)
```

Now, I'll plot just the lipid pairs

```{r}
#| dev: png

cluster <- c(
  "Cupuaçu-Cacau",
  "Bacuri-Pracaxi",
  "Cacau-Tucumã",
  "Cupuaçu-Tucumã",
  "Cupuaçu-Açaí",
  "Tucumã-Urucum",
  "Manga-Andiroba",
  "Murumuru-Jambu",
  "Manga-Pracaxi",
  "Murumuru-Cupuaçu",
  "Ucuuba-Andiroba"
)

coord <- as.data.frame(coefPca$x)

head(coefPca)

coord$lipid <- rownames(coord)

coord$group <- coord$lipid %in% cluster

groupCoord <- subset(coord, group)

hull <- groupCoord[
  chull(
    groupCoord$PC1,
    groupCoord$PC2
  ),
]



pcaInd <- fviz_pca_ind(
  X = coefPca,
  repel = TRUE,
  col.ind = "#FA5F55",
  #fill.ind = "contrib",
  pointsize = 2,
  arrowsize = 0.8
) + theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "PCA of Lipid Pairs"
  ) + geom_polygon(
    data = hull,
    aes(
      x = PC1,
      y = PC2
    ),
    inherit.aes = FALSE,
    fill = NA,
    color = "#40B5AD",
    linewidth = 0.7
  )
 # stat_ellipse(
  #  data = subset(coord, group),
   # aes(
    #  x = PC1,
    #  y = PC2
   # ),
   # linewidth = 1
 # )
  
pcaInd

ggsave(
  "pcaInd.png",
  plot = pcaInd,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 600
)
```

Now I shall plot the bidimensional graph with the parameter's vectors and lipid pair points.

```{r}
#| dev: png

pcaPlot <- fviz_pca(
  X = coefPca,
  repel = TRUE,
  col.var = "#40B5AD",
  col.ind = "#FA5F55",
  #fill.ind = "contrib",
  pointsize = 2,
  arrowsize = 0.8
) + theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "PCA of Scheffé model's coefficients"
  )

pcaPlot

ggsave(
  "pcaLipidPair.png",
  plot = pcaPlot,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 600
)


```

Now I'll plot the lipid pairs with their **Scheffé equation's degree fitted**.

```{r}
#| dev: png

pcaModel <- fviz_pca_ind(
  X = coefPca,
  geom = "point",
  repel = TRUE,
  col.ind = "#FA5F55",
  habillage = coefData$Model,
  mean.point = FALSE,
  alpha.ind = 0.7,
  #addEllipses = TRUE,
  #fill.ind = "contrib",
  pointsize = 5,
  arrowsize = 0.8
) + theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "PCA of Lipid Pairs by Degree of Scheffé's Equation"
  ) 

pcaModel

ggsave(
  "pcaModel.png",
  plot = pcaModel,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 600
)
```

Now, I'll plot the lipid pairs with their nature as butter or oil, to identify some relationship.

```{r}
#| dev: png

coefDataComp <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/Topicos_em_Nanotecnologia/TabelaCoeficientesNanoIACad.csv")

head(coefDataComp)

pcaType <- fviz_pca_ind(
  X = coefPca,
  geom = "point",
  repel = TRUE,
  col.ind = "#FA5F55",
  habillage = coefDataComp$Type,
  mean.point = FALSE,
  alpha.ind = 0.7,
 #addEllipses = TRUE,
  #fill.ind = "contrib",
  pointsize = 5,
  arrowsize = 0.8
) + theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "PCA of Lipid Pairs by Physic State"
  ) 

pcaType

ggsave(
  "pcaType.png",
  plot = pcaType,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 600
)
```

Now I'll plot with the fusion point:

```{r}
#| dev: png

#install.packages("ggrepel")
library(ggrepel)

pcaFP <- fviz_pca_ind(
  X = coefPca,
  repel = TRUE,
  col.ind = "#FA5F55",
  label = "none",
  #fill.ind = "contrib",
  pointsize = 2,
  arrowsize = 0.8
) + theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "PCA of Lipid Pairs by fusion point"
  )+
  geom_text_repel(
    aes(
      x = coefPca$x[,1],
      y = coefPca$x[,2],
      label = coefDataComp$FP
    )
  )

pcaFP

ggsave(
  "pcaFP.png",
  plot = pcaFP,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 600
)
```

now, I'll plot the **difference of fusion points**.

```{r}
#| dev: png

pcaDeltaFP <- fviz_pca_ind(
  X = coefPca,
  repel = TRUE,
  col.ind = "#FA5F55",
  label = "none",
  #fill.ind = "contrib",
  pointsize = 2,
  arrowsize = 0.8
) + theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = expression(
      bold(
        "PCA of Lipid Pairs by"~Delta~"Fusion Point"
        )
    )
  )+
  geom_text_repel(
    aes(
      x = coefPca$x[,1],
      y = coefPca$x[,2],
      label = coefDataComp$deltaFP
    )
  )

pcaDeltaFP

ggsave(
  "pcaDeltaFP.png",
  plot = pcaDeltaFP,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 600
)
```

Lets now plot the points with the **percentage of saturated fatty acids**:

```{r}
#| dev: png

pcaSfa <- fviz_pca_ind(
  X = coefPca,
  repel = TRUE,
  col.ind = "#FA5F55",
  label = "none",
  #fill.ind = "contrib",
  pointsize = 2,
  arrowsize = 0.8
) + theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "PCA of Lipid Pairs by percentage of saturated fatty acids"
  )+
  geom_text_repel(
    aes(
      x = coefPca$x[,1],
      y = coefPca$x[,2],
      label = coefDataComp$esfa
    )
  )

pcaSfa

ggsave(
  "pcaSfa.png",
  plot = pcaSfa,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 600
)
```

Now, I'll plot with the percentage of **unsaturated fatty acids**.

```{r}
#| dev: png

pcaUfa <- fviz_pca_ind(
  X = coefPca,
  repel = TRUE,
  col.ind = "#FA5F55",
  label = "none",
  #fill.ind = "contrib",
  pointsize = 2,
  arrowsize = 0.8
) + theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "PCA of Lipid Pairs by percentage of unsaturated fatty acids"
  )+
  geom_text_repel(
    aes(
      x = coefPca$x[,1],
      y = coefPca$x[,2],
      label = coefDataComp$eufa
    )
  )

pcaUfa

ggsave(
  "pcaUfa.png",
  plot = pcaUfa,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 600
)
```

Plotting now the **difference of the percentage of saturated fatty acids**.

```{r}
#| dev: png

clusterSfa <- c(
  "Ucuuba-Buriti",
  "Murumuru-Buriti",
  "Murumuru-Açaí"
)

coordSfa <- as.data.frame(coefPca$x)

head(coefPca)

coordSfa$lipid <- rownames(coordSfa)

coordSfa$group <- coordSfa$lipid %in% clusterSfa

groupCoordSfa <- subset(coordSfa, group)

hullSfa <- groupCoordSfa[
  chull(
    groupCoordSfa$PC1,
    groupCoordSfa$PC2
  ),
]



pcaDeltaSfa <- fviz_pca_ind(
  X = coefPca,
  repel = TRUE,
  col.ind = "#FA5F55",
  label = "none",
  #fill.ind = "contrib",
  pointsize = 2,
  arrowsize = 0.8
) + theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = expression(
      bold(
        "PCA of Lipid Pairs by"~Delta~"Percentage of Saturated Fatty Acids"
      )
    )
  )+
  geom_text_repel(
    aes(
      x = coefPca$x[,1],
      y = coefPca$x[,2],
      label = coefDataComp$deltaSfa
    )
  ) + geom_polygon(
    data = hullSfa,
    aes(
      x = PC1,
      y = PC2
    ),
    inherit.aes = FALSE,
    fill = NA,
    color = "#40B5AD",
    linewidth = 1.3
  )

pcaDeltaSfa

ggsave(
  "pcaDeltaSfa.png",
  plot = pcaDeltaSfa,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 600
)
```

Now, I'll plot the points with the label of **major fatty acid per lipid**:

```{r}
#| dev: png

pcaFA <- fviz_pca_ind(
  X = coefPca,
  geom = "point",
  pointshape = 16,
  repel = TRUE,
  col.ind = "#FA5F55",
  habillage = coefDataComp$FA,
  mean.point = FALSE,
  alpha.ind = 0.7,
  #addEllipses = TRUE,
  #fill.ind = "contrib",
  pointsize = 5,
  arrowsize = 0.8
) + theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "PCA of Lipid Pairs by major fatty acid",
    color = "Major fatty acids"
  ) 

pcaFA

ggsave(
  "pcaFA.png",
  plot = pcaFA,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 600
)
```

Now, I'll try to use *clustering* methods to separate the lipid pairs into **clusters** and try to extract some data.

```{r}
#| dev: png

#| message: FALSE

distPair <- dist(
  scale(coefDataPca)
)
distPair

fviz_dist(
  distPair,
  gradient = list(low = "blue",
                  mid = "white",
                  high = "red")
  )

coefHc <- hclust(
  distPair,
  method = "ward.D2"
)

coefHc

plot(coefHc)

pairTree <- cutree(
  coefHc,
  k=3
)

pairTree

pairClut <- fviz_pca_ind(
  X = coefPca,
  geom = "point",
  habillage = factor(pairTree),
  addEllipses = TRUE,
  mean.point = FALSE,
  pointsize = 5,
  alpha.ind = 0.5
  #title = "Lipid Pairs by cluster"
) + theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "Lipid Pairs by Cluster"
    #color = "Clusters"
  ) 

pairClut

ggsave(
  "pairClut.png",
  plot = pairClut,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 600
)

head(coefDataPca)

fviz_dist(
  distPair
)

coefEhc <- eclust(
  x = coefDataPca,
  FUNcluster = "hclust",
  K = 3,
  k.max = 3
)

fviz_cluster(
  coefEhc,
  geom = "point",
  repel = TRUE)

fviz_dend(
  x = coefEhc,
  lwd = 
)

library(ggdendro)
library(ggplot2)

coefDend <- dendro_data(coefEhc)

ggplot(
  segment(coefDend)
) +
  geom_segment(
    aes(
      x = x,
      y = y,
      xend = xend,
      yend = yend
    )
  ) #+
  #theme_classic()

#dendPlot

library(dendextend)

coefDend <- as.dendrogram(coefHc)

coefDend <- color_branches(
  coefDend,
  k = 3
)

png(
  "cluster.png",
  width = 18,
  height = 14,
  units = "cm",
  res = 600
)

plot(
  coefDend,
  main = "Cluster Dendrogram",
  xlab = "",
  sub = "",
  cex = 1.2
)

dev.off()
```
