---
title: "Lipids isolated effects on coefficients"
author: "Luís Romera"
format: html
editor: visual
---

```{r}
getwd()
```

In this document, I'm going to calculate the mean of the coefficients of each lipid used in the nanostructure formulations.

## Lipids:

- **Bacuri**

- **Açaí**

- **Pracaxi**

- **Cacau**

- **Tucumã**

- **Cupuaçu**

- **Pequi**

- **Manga**

- **Andiroba**

- **Murumuru**

- **Buriti**

- **Jambu**

- **Urucum**

- **Ucuuba**

I'm going to apply multiple linear regression to determine how each lipid dictates the **Scheffé's coefficients**.

```{r}
coefData <- read.csv("Tabela de coeficientes NanoIA.csv")
```

Let's fit the linear models

## B1

```{r}
B1 <- lm(
  b1 ~
    Bacuri +
    Açaí +
    Pracaxi +
    Cacau +
    Tucumã +
    Cupuaçu +
    Pequi +
    Manga +
    Andiroba +
    Murumuru +
    Buriti +
    Jambu +
    Urucum +
    Ucuuba,
  data = coefData
)

B1

summary(B1)

anova(B1)
```

## B2

```{r}

lm(
  b2 ~
    Bacuri +
    Açaí +
    Pracaxi +
    Cacau +
    Tucumã +
    Cupuaçu +
    Pequi +
    Manga +
    Andiroba +
    Murumuru +
    Buriti +
    Jambu +
    Urucum +
    Ucuuba,
  data = coefData
)
```

## B3

```{r}
lm(
  b3 ~
    Bacuri +
    Açaí +
    Pracaxi +
    Cacau +
    Tucumã +
    Cupuaçu +
    Pequi +
    Manga +
    Andiroba +
    Murumuru +
    Buriti +
    Jambu +
    Urucum +
    Ucuuba,
  data = coefData
)
```

## B12

```{r}
B12 <- lm(
  b12 ~
    Bacuri +
    Açaí +
    Pracaxi +
    Cacau +
    Tucumã +
    Cupuaçu +
    Pequi +
    Manga +
    Andiroba +
    Murumuru +
    Buriti +
    Jambu +
    Urucum +
    Ucuuba,
  data = coefData
)

B12

summary(B12)
```

## B13

```{r}

lm(
  b13~
    Bacuri +
    Açaí +
    Pracaxi +
    Cacau +
    Tucumã +
    Cupuaçu +
    Pequi +
    Manga +
    Andiroba +
    Murumuru +
    Buriti +
    Jambu +
    Urucum +
    Ucuuba,
  data = coefData
)
```

## B23

```{r}

lm(
  b23~
    Bacuri +
    Açaí +
    Pracaxi +
    Cacau +
    Tucumã +
    Cupuaçu +
    Pequi +
    Manga +
    Andiroba +
    Murumuru +
    Buriti +
    Jambu +
    Urucum +
    Ucuuba,
  data = coefData
)
```

Now, I'm going to apply PCA to the lipids alone.

```{r}

#| warning: False
#| message: FALSE

getwd()

coefSing <- read.csv("coeficientesSingular.csv")

coefSing <- coefSing[rowSums(is.na(coefSing)) < ncol(coefSing), ]

coefSing <- coefSing[, colSums(is.na(coefSing)) < nrow(coefSing)]

coefSing <- coefSing[!is.na(coefSing$Lipid), ]

coefSing <- coefSing[coefSing$Lipid != "", ]

nrow(coefSing)

str(coefSing)

head(coefSing)

pcaData <- coefSing[, c(
"b1",
"b2",
"b3",
"b12",
"b13",
"b23"
)]

rownames(pcaData) <- coefSing$Lipid

View(coefSing)

pca <- prcomp(
  pcaData,
  retx = TRUE,
  center = TRUE,
  scale. = TRUE
)
pca

summary(pca)

library(factoextra)

pcaEigLip <- fviz_eig(pca,
                      barcolor = "black",
                      addlabels = TRUE,
                      ggtheme = theme_minimal(),
                      #  parallel = TRUE 
                      )
pcaEigLip

ggsave(
  "pcaEigLip.png",
  plot = pcaEigLip,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 300
)

pcaLip <- fviz_pca_ind(
  X = pca,
  repel = TRUE,
  pointsize = 2
) + theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "PCA of lipids"
  )

pcaLip

ggsave(
  "pcaLip.png",
  plot = pcaLip,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 300
)

pcaBiplot <- fviz_pca(
  X = pca,
  repel = TRUE,
  #col.var = "#40B5AD",
  #col.ind = "#FA5F55",
  #fill.ind = "contrib",
  pointsize = 2,
  arrowsize = 0.8
) + theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "PCA of coefficients and lipids"
  )
pcaBiplot

ggsave(
  "pcaBiplot.png",
  plot = pcaBiplot,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 300
)

pcaVar <- fviz_pca_var(
  X = pca,
repel = TRUE,
arrowsize = 0.8,
col.var = "contrib",
gradient.cols = c("#40B5AD", "#FA5F55")
) +
    theme(
      plot.title = element_text(
        hjust = 0.5,
        face = "bold",
        size = 12
      ),
      axis.title = element_text(
        face = "bold",
        size = 10
      ),
      axis.text = element_text(size = 8)
    ) +
    labs(
      title = "PCA of coefficients"
    )
pcaVar

ggsave(
  "pcaVar.png",
  plot = pcaVar,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 300
)

round(pca$rotation, 3)
```

## Clustering

```{r}

distMat <- dist(scale(pcaData))

hc <- hclust(
  distMat,
  method = "ward.D2"
)

hc

cluster <- cutree(
  hc,
  k = 3
)
cluster

pcaHcLipid <- fviz_pca_ind(
  pca,
  geom = "point",
  habillage = factor(cluster),
  pointshape = 16,
  #addEllipses = TRUE,
  fill.ind = "black",
  mean.point = FALSE,
  ellipse.level = 0.95,
  pointsize = 4,
  repel = TRUE
) +
  theme_classic()

pcaHcLipid

ggsave(
  "pcaHcLipid.png",
  plot = pcaHcLipid,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 300
)

pcaComp <- fviz_pca_biplot(
  pca,
  habillage = factor(cluster),
  addEllipses = TRUE,
  ellipse.level = 0.95,
  pointsize = 4,
  repel = TRUE
) 

#pcaComp

#searching the optimal number of clusters

fviz_nbclust(
  scale(pcaData),
  FUNcluster = hcut,
  method = "silhouette",
  k.max = 10,
  print.summary = FALSE
)

pcaEclust <- eclust(
  x = scale(pcaData),
  FUNcluster = "hclust",
  k = 4,
  k.max = 10
  #graph = TRUE
)

pcaEclust

summary(pcaEclust)

cutree(
  pcaEclust,
  k = 5
)

pcaCluster <- fviz_cluster(
    pcaEclust,
    #repel = TRUE,
    show.clust.cent = FALSE,
    pointsize = 4,
    palette = c("#FA6F55", "#339E2C", "#CCC941", "#418DBA"),
  ) + theme_classic() +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "Cluster Plot of Lipids"
  )

pcaCluster

ggsave(
  "pcaCluster.png",
  plot = pcaCluster,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 300
)

identical(
  pcaEclust$cluster,
  cutree(pcaEclust,k=3)
)

hc <- hclust(
  dist(
    scale(pcaData)
    ),
  method = "ward.D2"
)

cluster1 <- cutree(
  hc,
  k = 3
)

cluster2 <- pcaEclust$cluster

all(cluster1 == cluster2)

cluster2
```

Now, I'll try to plot with the physical state label

```{r}

lipidComp <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/Topicos_em_Nanotecnologia/coeficientesSingularComp.csv")


library(ggrepel)

  pcaClusterType <- fviz_cluster(
    pcaEclust,
    geom = "point",
    show.clust.cent = FALSE,
    pointsize = 4,
    palette = c("#FA6F55", "#339E2C", "#CCC941", "#418DBA"),
    
  ) +
    theme_classic() +
    geom_text_repel(
      aes(
        label = lipidComp$type
      ),
      size = 4
    ) + 
    theme(
      plot.title = element_text(
        hjust = 0.5,
        face = "bold",
        size = 12
      ),
      axis.title = element_text(
        face = "bold",
        size = 10
      ),
      axis.text = element_text(size = 8)
    ) +
    labs(
      title = "Cluster Plot with Physical state"
    )
  
  pcaClusterType


  ggsave(
    "pcaClusterType.png",
    plot = pcaClusterType,
    width = 18,
    height = 14,
    units = "cm",
    dpi = 300
  )
```

Now I'll plot **fusion point**

```{r}

pcaClusterFP <- fviz_cluster(
  pcaEclust,
  geom = "point",
  show.clust.cent = FALSE,
  pointsize = 4,
  palette = c("#FA6F55", "#339E2C", "#CCC941", "#418DBA"),
) +
  theme_classic() +
  geom_text_repel(
    aes(
      label = lipidComp$FP
    )
  ) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "Cluster Plot with Fusion Point"
  )

pcaClusterFP


ggsave(
  "pcaClusterFP.png",
  plot = pcaClusterFP,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 300
)
```

Now, I'll plot with the **percentage of unsatyrated fatty acids**

```{r}

pcaClusterUFA <- fviz_cluster(
  pcaEclust,
  geom = "point",
  show.clust.cent = FALSE,
  pointsize = 4,
  palette = c("#FA6F55", "#339E2C", "#CCC941", "#418DBA"),
) +
  theme_classic() +
  geom_text_repel(
    aes(
      label = lipidComp$UFA
    )
  ) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "Cluster Plot with Percentage of Unsaturated Fatty Acids"
  )

pcaClusterUFA


ggsave(
  "pcaClusterUFA.png",
  plot = pcaClusterUFA,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 300
)
```

Now with **major fatty acid chain**

```{r}

pcaClusterFA <- fviz_cluster(
  pcaEclust,
  geom = "point",
  show.clust.cent = FALSE,
  pointsize = 4,
  palette = c("#FA6F55", "#339E2C", "#CCC941", "#418DBA"),
) +
  theme_classic() +
  geom_text_repel(
    aes(
      label = lipidComp$FA
    )
  ) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 12
    ),
    axis.title = element_text(
      face = "bold",
      size = 10
    ),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "Cluster Plot with major Fatty Acid chain"
  )

pcaClusterFA


ggsave(
  "pcaClusterFA.png",
  plot = pcaClusterFA,
  width = 18,
  height = 14,
  units = "cm",
  dpi = 300
)
```

## Dendrogram

```{r}

#library(factoextra)
library(dendextend)

dend <- as.dendrogram(hc)

dend <- color_branches(
  dend,
  k = 4
)

dev.new()

plot(dend)

abline(
  h = 4,
  col = "black",
  lwd = 2,
  lty = 2
)

class(dend)

dev.cur()
#png(
 # "dend.png",
 # width = 18,
#  height = 14,
 # units = "cm",
#  res = 300
#)

plot(
  dend,
  main = "Cluster Dendrogram",
  xlab = "",
  sub = "",
  cex = 1.2
)

dev.list()

dev.off()
```

## Heatmap

```{r}

library(pheatmap)

pcaDataScale <- scale(pcaData)

heatMap <- pheatmap(
  pcaDataScale,
  clustering_method = "ward.D2",
  border_color = NA,
  fontsize_row = 12,
  fontsize_col = 12,
  main = "Lipid clustering by Scheffé's Coefficients"
)

class(heatMap)

heatMap
```
