---
title: "Nano IA Mixture Designs"
author: "Luís Felipe Romera"
format: html
editor: visual
---

In this document, I'm going to perform **mixture designs** on the formulation data of the different *nanoparticles* based on amazonian lipids.

First I'm going to explore the **PDI** measurement, than I'm going to explore the nanoparticle's **size**.

# PDI

## Nano Bacuri-Açaí

```{r}
#| message: false
#| warning: false

setwd("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas")

bacuriAcai <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas/NanoBacuriAcai.csv")

head(bacuriAcai)

library(mixexp)

bacuriAcaiVar <- c("Bacuri", "Acai", "Brijo10")

bacuriAcaiMod1 <- MixModel(bacuriAcai, "PDI", bacuriAcaiVar, 1)

summary(bacuriAcaiMod1)

anova(bacuriAcaiMod1)

bacuriAcaiMod2 <- MixModel(bacuriAcai, "PDI", bacuriAcaiVar, 2)

summary(bacuriAcaiMod2)

anova(bacuriAcaiMod2)

bacuriAcaiMod3 <- MixModel(bacuriAcai, "PDI", bacuriAcaiVar, 4)

summary(bacuriAcaiMod3)

anova(bacuriAcaiMod3)

anova(bacuriAcaiMod1, bacuriAcaiMod2)

anova(bacuriAcaiMod1, bacuriAcaiMod3)

summary(bacuriAcaiMod1)

par(mfrow = c(2,2))

plot(bacuriAcaiMod1)
```

### Results PDI Bacuri-Açaí

The model that fitted the PDI data the best was the linear model. Bacuri is the only variable to have a significant effect on the model. With a positive coefficient of 2.34 on an Adjusted R-Squared of 0.93.

## Nano Bacuri-Pracaxi

```{r}
bacuriPracaxi <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas/NanoBacuriPracaxi.csv")

head(bacuriPracaxi)

bacuriPracaxiVar <-  c("Bacuri", "Pracaxi", "Brijo10")

bacuriPracaxiMod1 <- MixModel(bacuriPracaxi, "PDI", bacuriPracaxiVar, 1)

summary(bacuriPracaxiMod1)

anova(bacuriPracaxiMod1)

bacuriPracaxiMod2 <- MixModel(bacuriPracaxi, "PDI", bacuriPracaxiVar, 2)

summary(bacuriPracaxiMod2)

anova(bacuriPracaxiMod2)

bacuriPracaxiMod3 <- MixModel(bacuriPracaxi, "PDI", bacuriPracaxiVar, 3)

summary(bacuriPracaxiMod3)

anova(bacuriPracaxiMod3)

anova(bacuriPracaxiMod1,bacuriPracaxiMod2)

anova(bacuriPracaxiMod1, bacuriPracaxiMod3)

summary(bacuriPracaxiMod1)

par(mfrow = c(2, 2))

plot(bacuriAcaiMod1)
```

### Results PDI Nano Bacuri-Pracaxi

The model that best fit the data is the linear model, with Brij O10 being the only parameter that contributes individually for the model (Pr(t) \< 0.05), although both Brij O10 and Bacuri are paramenters that explain the data better than a reduced model (Pr(F) \< 0.05), with an estimate of 1.57 and -0.1, respectively.

## Nano Cacau-Tucumã

```{r}
#| warning: false
#| message: false

cacauTucuma <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas/NanoCacauTucuma.csv")

head(cacauTucuma)

cacauTucumaVar <- c("Cacau", "Tucuma", "Brijo10")

cacauTucumaMod1 <- MixModel(cacauTucuma, "PDI", cacauTucumaVar, 1)

summary(cacauTucumaMod1)

anova(cacauTucumaMod1)

cacauTucumaMod2 <- MixModel(cacauTucuma, "PDI", cacauTucumaVar, 2)

summary(cacauTucumaMod2)

anova(cacauTucumaMod1)

cacauTucumaMod3 <- MixModel(cacauTucuma, "PDI", cacauTucumaVar, 3)

summary(cacauTucumaMod3)

anova(cacauTucumaMod3)

anova(cacauTucumaMod1, cacauTucumaMod2)

anova(cacauTucumaMod1, cacauTucumaMod3)

par(mfrow = c(2, 2))

plot(cacauTucumaMod1)
```

### Results PDI Nano Cacau-Tucumã

No model fitted the data well (best determinant coefficient = 0.06)

## Nano Cupuaçu-Açaí

```{r}
#| warning: false
#| message: false

cupuacuAcai <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas/NanoCupuacuAcai.csv")

head(cupuacuAcai)

cupuacuAcaiVar <- c("Cupuacu", "Acai", "Brijo10")

cupuacuAcaiMod1 <- MixModel(cupuacuAcai, "PDI", cupuacuAcaiVar, 1)

summary(cupuacuAcaiMod1)

anova(cupuacuAcaiMod1)

cupuacuAcaiMod2 <- MixModel(cupuacuAcai, "PDI", cupuacuAcaiVar, 2)

summary(cupuacuAcaiMod2)

anova(cupuacuAcaiMod2)

cupuacuAcaiMod3 <- MixModel(cupuacuAcai, "PDI", cupuacuAcaiVar, 3)

summary(cupuacuAcaiMod3)

anova(cupuacuAcaiMod3)

anova(cupuacuAcaiMod1, cupuacuAcaiMod2)

anova(cupuacuAcaiMod1, cupuacuAcaiMod3)

par(mfrow = c(2, 2))

plot(cupuacuAcaiMod1)
```

### Results PDI Nano Cupuaçu-Açaí

The model that best fitted the data was the linear model. Açaí was the only parameter that explains the variance alone, although it, along Cupuaçu, explains the variance better than a simpler model, with an estimate of 1.16 and 0.7, respectively.

## Nano Cupuaçu-Cacau

```{r}
#| warning: false
#| message: false

cupuacuCacau <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas/NanoCupuacuCacau.csv")

head(cupuacuCacau)

cupuacuCacauVar <- c("Cupuacu", "Cacau", "Brijo10")

cupuacuCacauMod1 <- MixModel(cupuacuCacau, "PDI", cupuacuCacauVar, 1)

summary(cupuacuCacauMod1)

anova(cupuacuCacauMod1)

cupuacuCacauMod2 <- MixModel(cupuacuCacau, "PDI", cupuacuCacauVar, 2)

summary(cupuacuCacauMod2)

anova(cupuacuCacauMod2)

anova(cupuacuCacauMod1, cupuacuCacauMod2)

par(mfrow = c(2, 2))

plot(cupuacuCacauMod1)
```

### Results PDI Cupuaçu-Cacau

The model that fitted the data the best was the linear model, with a determinant coefficient of 0.75 and with Cupuaçu being the only parameter that contributes alone to the model and that explains the variance better than a reduced model.

## Nano Cupuaçu-Pequi

```{r}
#| warning: false
#| message: false

cupuacuPequi <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas/NanoCupuacuPequi.csv")

head(cupuacuPequi)

cupuacuPequiVar <- c("Cupuacu", "Pequi", "Brijo10")

cupuacuPequiMod1 <- MixModel(cupuacuPequi, "PDI", cupuacuPequiVar, 1)

summary(cupuacuPequiMod1)

anova(cupuacuPequiMod1)

cupuacuPequiMod2 <- MixModel(cupuacuPequi, "PDI", cupuacuPequiVar, 2)

summary(cupuacuPequiMod2)

anova(cupuacuPequiMod2)

cupuacuPequiMod3 <- MixModel(cupuacuPequi, "PDI", cupuacuPequiVar, 4)

summary(cupuacuPequiMod3)

anova(cupuacuPequiMod3)

anova(cupuacuPequiMod1, cupuacuPequiMod2)

anova(cupuacuPequiMod1, cupuacuPequiMod3)

par(mfrow = c(2, 2))

plot(cupuacuPequiMod2)

cupuacuPequi$ponto <- interaction(
  cupuacuPequi$Cupuacu,
  cupuacuPequi$Pequi,
  cupuacuPequi$Brijo10
)

modScheffe <- lm(
  PDI ~ Cupuacu + Pequi + Brijo10 +
    Cupuacu:Pequi + Cupuacu:Brijo10 + Pequi:Brijo10 - 1,
  data = cupuacuPequi
)

mod_sat <- lm(
  PDI ~ ponto,
  data = cupuacuPequi
)


#Lack of Fit test
anova(modScheffe, mod_sat)
```

### Results PDI Cupuaçu-Pequi

The quadratic model was chosen to fit the data with an **adjusted R² = 0.91**, and with the interaction of Cupuaçuu and Pequi with an estimator of **- 3.91**.

## Nano Cupuaçu-Tucumã

```{r}

#| warning: false
#| message: false

cupuacuTucuma <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas/NanoCupuacuTucuma.csv")

head(cupuacuTucuma)

cupuacuTucumaVar <- c("Cupuacu", "Tucuma", "Brijo10")

cupuacuTucumaMod1 <- MixModel(cupuacuTucuma, "PDI", cupuacuTucumaVar, 1)

summary(cupuacuTucumaMod1)

anova(cupuacuTucumaMod1)

cupuacuTucumaMod2 <- MixModel(cupuacuTucuma, "PDI", cupuacuTucumaVar, 2)

summary(cupuacuTucumaMod2)

anova(cupuacuTucumaMod2)

#cupuacuPequiMod3 <- MixModel(cupuacuPequi, "PDI", cupuacuPequiVar, 4)

#summary(cupuacuPequiMod3)

#anova(cupuacuPequiMod3)

anova(cupuacuTucumaMod1, cupuacuTucumaMod2)

#anova(cupuacuPequiMod1, cupuacuPequiMod3)

par(mfrow = c(2, 2))

plot(cupuacuTucumaMod1)

cupuacuTucuma$ponto <- interaction(
  cupuacuTucuma$Cupuacu,
  cupuacuTucuma$Tucuma,
  cupuacuTucuma$Brijo10
)

#modScheffe <- lm(
#  PDI ~ Cupuacu + Pequi + Brijo10 +
#    Cupuacu:Pequi + Cupuacu:Brijo10 + Pequi:Brijo10 - 1,
#  data = cupuacuPequi
#)

modSat <- lm(
  PDI ~ ponto,
  data = cupuacuTucuma
)


#Lack of Fit test
anova(cupuacuTucumaMod1, modSat)

View(cupuacuTucuma)
```

### Results

The linear model was chosen to fit the data

## Nano Manga-Andiroba

```{r}
#| warning: false
#| message: false

mangaAndiroba <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas/NanoMangaAndiroba.csv")

head(mangaAndiroba)

mangaAndirobaVar <- c("Manga", "Andiroba", "Brijo10")

mangaAndirobaMod1 <- MixModel(mangaAndiroba, "PDI", mangaAndirobaVar, 1)

summary(mangaAndirobaMod1)

anova(mangaAndirobaMod1)

mangaAndirobaMod2 <- MixModel(mangaAndiroba, "PDI", mangaAndirobaVar, 2)

summary(mangaAndirobaMod2)

anova(mangaAndirobaMod2)

#cupuacuPequiMod3 <- MixModel(cupuacuPequi, "PDI", cupuacuPequiVar, 4)

#summary(cupuacuPequiMod3)

#anova(cupuacuPequiMod3)

anova(mangaAndirobaMod1, mangaAndirobaMod2)

#anova(cupuacuPequiMod1, cupuacuPequiMod3)

par(mfrow = c(2, 2))

plot(mangaAndirobaMod1)

mangaAndiroba$ponto <- interaction(
  mangaAndiroba$Manga,
  mangaAndiroba$Andiroba,
  mangaAndiroba$Brijo10
)

#modScheffe <- lm(
#  PDI ~ Cupuacu + Pequi + Brijo10 +
#    Cupuacu:Pequi + Cupuacu:Brijo10 + Pequi:Brijo10 - 1,
#  data = cupuacuPequi
#)

modSat <- lm(
  PDI ~ ponto,
  data = mangaAndiroba
)


#Lack of Fit test
anova(mangaAndirobaMod1, modSat)

View(cupuacuTucuma)
```

### Results

The **Linear Model** was chosen to fit the data.

## Nano Manga-Pracaxi

```{r}
#| warning: false
#| message: false

mangaPracaxi <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas/NanoMangaPracaxi.csv")

#head(mangaAndiroba)

mangaPracaxiVar <- c("Manga", "Pracaxi", "Brijo10")

mangaPracaxiMod1 <- MixModel(mangaPracaxi, "PDI", mangaPracaxiVar, 1)

summary(mangaPracaxiMod1)

anova(mangaPracaxiMod1)

mangaPracaxiMod2 <- MixModel(mangaPracaxi, "PDI", mangaPracaxiVar, 2)

summary(mangaPracaxiMod2)

anova(mangaPracaxiMod2)

#cupuacuPequiMod3 <- MixModel(cupuacuPequi, "PDI", cupuacuPequiVar, 4)

#summary(cupuacuPequiMod3)

#anova(cupuacuPequiMod3)

anova(mangaPracaxiMod1, mangaPracaxiMod2)

#anova(cupuacuPequiMod1, cupuacuPequiMod3)

par(mfrow = c(2, 2))

plot(mangaPracaxiMod1)

mangaPracaxi$ponto <- interaction(
  mangaPracaxi$Manga,
  mangaPracaxi$Pracaxi,
  mangaPracaxi$Brijo10
)

#modScheffe <- lm(
#  PDI ~ Cupuacu + Pequi + Brijo10 +
#    Cupuacu:Pequi + Cupuacu:Brijo10 + Pequi:Brijo10 - 1,
#  data = cupuacuPequi
#)

modSat <- lm(
  PDI ~ ponto,
  data = mangaPracaxi
)


#Lack of Fit test
anova(mangaAndirobaMod1, modSat)

#View(cupuacuTucuma)
```

### Results

The **Linear Model** was chosen to fit the data.

## Nano Murumuru-Açaí

```{r}
#| warning: false
#| message: false

murumuruAcai <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas/NanoMurumuruAcai.csv")

#head(mangaAndiroba)

murumuruAcaiVar <- c("Murumuru", "Acai", "Brijo10")

murumuruAcaiMod1 <- MixModel(murumuruAcai, "PDI", murumuruAcaiVar, 1)

summary(murumuruAcaiMod1)

anova(murumuruAcaiMod1)

murumuruAcaiMod2 <- MixModel(murumuruAcai, "PDI", murumuruAcaiVar, 2)

summary(murumuruAcaiMod2)

anova(murumuruAcaiMod2)

#cupuacuPequiMod3 <- MixModel(cupuacuPequi, "PDI", cupuacuPequiVar, 4)

#summary(cupuacuPequiMod3)

#anova(cupuacuPequiMod3)

anova(murumuruAcaiMod1, murumuruAcaiMod2)

#anova(cupuacuPequiMod1, cupuacuPequiMod3)

par(mfrow = c(2, 2))

plot(murumuruAcaiMod2)

murumuruAcai$ponto <- interaction(
  murumuruAcai$Murumuru,
  murumuruAcai$Acai,
  murumuruAcai$Brijo10
)

modScheffe <- lm(
  PDI ~ Murumuru + Acai + Brijo10 +
    Murumuru:Acai + Murumuru:Brijo10 + Acai:Brijo10 - 1,
  data = murumuruAcai
)

modSat <- lm(
  PDI ~ ponto,
  data = murumuruAcai
)


#Lack of Fit test
anova(modScheffe, modSat)

View(murumuruAcai)
```

### Results

We were not able to fit a model.

## Nano Murumuru-Buriti

```{r}
#| warning: false
#| message: false

murumuruBuriti <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas/NanoMurumuruBuriti.csv")

#head(mangaAndiroba)

murumuruBuritiVar <- c("Murumuru", "Buriti", "Brijo10")

murumuruBuritiMod1 <- MixModel(murumuruBuriti, "PDI", murumuruBuritiVar, 1)

summary(murumuruBuritiMod1)

anova(murumuruBuritiMod1)

murumuruBuritiMod2 <- MixModel(murumuruBuriti, "PDI", murumuruBuritiVar, 2)

summary(murumuruBuritiMod2)

anova(murumuruBuritiMod2)

#cupuacuPequiMod3 <- MixModel(cupuacuPequi, "PDI", cupuacuPequiVar, 4)

#summary(cupuacuPequiMod3)

#anova(cupuacuPequiMod3)

anova(murumuruBuritiMod1, murumuruBuritiMod2)

#anova(cupuacuPequiMod1, cupuacuPequiMod3)

par(mfrow = c(2, 2))

plot(murumuruAcaiMod2)

murumuruBuriti$ponto <- interaction(
  murumuruBuriti$Murumuru,
  murumuruBuriti$Buriti,
  murumuruBuriti$Brijo10
)

modScheffe <- lm(
  PDI ~ Murumuru + Acai + Brijo10 +
    Murumuru:Acai + Murumuru:Brijo10 + Acai:Brijo10 - 1,
  data = murumuruAcai
)

modSat <- lm(
  PDI ~ ponto,
  data = murumuruBuriti
)


#Lack of Fit test
anova(murumuruBuritiMod1, modSat)

View(murumuruBuriti)
```

### Results

No model was fitted to the data.

## Nano Murumuru-Cupuaçu

```{r}
#| warning: false
#| message: false

murumuruCupuacu <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas/NanoMurumuruCupuacu.csv")

head(murumuruCupuacu)

murumuruCupuacu$Cupuacu <- murumuruCupuacu$Cupuaçu

murumuruCupuacuVar <- c("Murumuru", "Cupuacu", "Brijo10")

murumuruCupuacuMod1 <- MixModel(murumuruCupuacu, "PDI", murumuruCupuacuVar, 1)

summary(murumuruCupuacuMod1)

anova(murumuruCupuacuMod1)

murumuruCupuacuMod2 <- MixModel(murumuruCupuacu, "PDI", murumuruCupuacuVar, 2)

summary(murumuruCupuacuMod2)

anova(murumuruCupuacuMod2)

#cupuacuPequiMod3 <- MixModel(cupuacuPequi, "PDI", cupuacuPequiVar, 4)

#summary(cupuacuPequiMod3)

#anova(cupuacuPequiMod3)

anova(murumuruCupuacuMod1, murumuruCupuacuMod2)

#anova(cupuacuPequiMod1, cupuacuPequiMod3)

par(mfrow = c(2, 2))

plot(murumuruAcaiMod2)

murumuruCupuacu$ponto <- interaction(
  murumuruCupuacu$Murumuru,
  murumuruCupuacu$Cupuacu,
  murumuruCupuacu$Brijo10
)

modScheffe <- lm(
  PDI ~ Murumuru + Acai + Brijo10 +
    Murumuru:Acai + Murumuru:Brijo10 + Acai:Brijo10 - 1,
  data = murumuruAcai
)

modSat <- lm(
  PDI ~ ponto,
  data = murumuruCupuacu
)


#Lack of Fit test
anova(murumuruCupuacuMod1, modSat)

View(murumuruBuriti)
```

### Results

The **Quadratic model** was chosen to fit the data.

## Nano Murumuru-Jambu

```{r}
#| warning: false
#| message: false

murumuruJambu <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas/NanoMurumuruJambu.csv")

#head(murumuruJambu)

#murumuruCupuacu$Cupuacu <- murumuruCupuacu$Cupuaçu

murumuruJambuVar <- c("Murumuru", "Jambu", "Brijo10")

murumuruJambuMod1 <- MixModel(murumuruJambu, "PDI", murumuruJambuVar, 1)

summary(murumuruJambuMod1)

anova(murumuruJambuMod1)

murumuruJambuMod2 <- MixModel(murumuruJambu, "PDI", murumuruJambuVar, 2)

summary(murumuruJambuMod2)

anova(murumuruJambuMod2)

#cupuacuPequiMod3 <- MixModel(cupuacuPequi, "PDI", cupuacuPequiVar, 4)

#summary(cupuacuPequiMod3)

#anova(cupuacuPequiMod3)

anova(murumuruJambuMod1, murumuruJambuMod2)

#anova(cupuacuPequiMod1, cupuacuPequiMod3)

par(mfrow = c(2, 2))

plot(murumuruAcaiMod2)

murumuruJambu$ponto <- interaction(
  murumuruJambu$Murumuru,
  murumuruJambu$Jambu,
  murumuruJambu$Brijo10
)

modScheffe <- lm(
  PDI ~ Murumuru + Acai + Brijo10 +
    Murumuru:Acai + Murumuru:Brijo10 + Acai:Brijo10 - 1,
  data = murumuruAcai
)

modSat <- lm(
  PDI ~ ponto,
  data = murumuruJambu
)


#Lack of Fit test
anova(murumuruJambuMod1, modSat)

#View(murumuruBuriti)
```

### Results

The **Linear model** was chosen to fit the data.

## Nano Tucumã-Tucumã

```{r}
#| warning: false
#| message: false

tucumaTucuma <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas/NanoTucumaTucuma.csv")

head(tucumaTucuma)

#murumuruCupuacu$Cupuacu <- murumuruCupuacu$Cupuaçu

tucumaTucumaVar <- c("Tucuma", "Tucuma.1", "Brijo10")

tucumaTucumaMod1 <- MixModel(tucumaTucuma, "PDI", tucumaTucumaVar, 1)

summary(tucumaTucumaMod1)

anova(tucumaTucumaMod1)

tucumaTucumaMod2 <- MixModel(tucumaTucuma, "PDI", tucumaTucumaVar, 2)

summary(tucumaTucumaMod2)

anova(tucumaTucumaMod2)

#cupuacuPequiMod3 <- MixModel(cupuacuPequi, "PDI", cupuacuPequiVar, 4)

#summary(cupuacuPequiMod3)

#anova(cupuacuPequiMod3)

anova(tucumaTucumaMod1, tucumaTucumaMod2)

#anova(cupuacuPequiMod1, cupuacuPequiMod3)

par(mfrow = c(2, 2))

plot(tucumaTucumaMod2)

tucumaTucuma$ponto <- interaction(
  tucumaTucuma$Tucuma,
  tucumaTucuma$Tucuma.1,
  tucumaTucuma$Brijo10
)

modScheffe <- lm(
  PDI ~ Murumuru + Acai + Brijo10 +
    Murumuru:Acai + Murumuru:Brijo10 + Acai:Brijo10 - 1,
  data = murumuruAcai
)

modSat <- lm(
  PDI ~ ponto,
  data = tucumaTucuma
)


#Lack of Fit test
anova(tucumaTucumaMod2, modSat)

#View(murumuruBuriti)
```

### Results

The **Quadratic Model** was chosen to fit the data.

## Nano Tucumã-Urucum

```{r}
#| warning: false
#| message: false

tucumaUrucum <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas/NanoTucumaUrucum.csv")

#head(tucumaTucuma)

#murumuruCupuacu$Cupuacu <- murumuruCupuacu$Cupuaçu

tucumaUrucumVar <- c("Tucuma", "Urucum", "Brijo10")

tucumaUrucumMod1 <- MixModel(tucumaUrucum, "PDI", tucumaUrucumVar, 1)

summary(tucumaUrucumMod1)

anova(tucumaUrucumMod1)

tucumaUrucumMod2 <- MixModel(tucumaUrucum, "PDI", tucumaUrucumVar, 2)

summary(tucumaUrucumMod2)

anova(tucumaUrucumMod2)

tucumaUrucumMod3 <- MixModel(tucumaUrucum, "PDI", tucumaUrucumVar, 4)

#summary(cupuacuPequiMod3)

#anova(cupuacuPequiMod3)

anova(tucumaUrucumMod1, tucumaUrucumMod2)

#anova(cupuacuPequiMod1, cupuacuPequiMod3)

par(mfrow = c(2, 2))

plot(tucumaTucumaMod2)

tucumaTucuma$ponto <- interaction(
  tucumaTucuma$Tucuma,
  tucumaTucuma$Tucuma.1,
  tucumaTucuma$Brijo10
)

modScheffe <- lm(
  PDI ~ Murumuru + Acai + Brijo10 +
    Murumuru:Acai + Murumuru:Brijo10 + Acai:Brijo10 - 1,
  data = murumuruAcai
)

modSat <- lm(
  PDI ~ ponto,
  data = tucumaTucuma
)


#Lack of Fit test
anova(tucumaTucumaMod2, modSat)

#View(murumuruBuriti)
```

### Results

The **Linear Model** was chosen to fit the data

## Nano Ucuuba-Andiroba

```{r}
#| warning: false
#| message: false

ucuubaAndiroba <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas/NanoUcuubaAndiroba.csv")

head(ucuubaAndiroba)

ucuubaAndiroba$Andiroba <- ucuubaAndiroba$Pracaxi

ucuubaAndirobaVar <- c("Ucuuba", "Andiroba", "Brijo10")

ucuubaAndirobaMod1 <- MixModel(ucuubaAndiroba, "PDI", ucuubaAndirobaVar, 1)

summary(ucuubaAndirobaMod1)

anova(ucuubaAndirobaMod1)

ucuubaAndirobaMod2 <- MixModel(ucuubaAndiroba, "PDI", ucuubaAndirobaVar, 2)

summary(ucuubaAndirobaMod2)

anova(ucuubaAndirobaMod2)

#tucumaUrucumMod3 <- MixModel(tucumaUrucum, "PDI", tucumaUrucumVar, 4)

#summary(cupuacuPequiMod3)

#anova(cupuacuPequiMod3)

anova(ucuubaAndirobaMod1, ucuubaAndirobaMod2)

#anova(cupuacuPequiMod1, cupuacuPequiMod3)

par(mfrow = c(2, 2))

#plot(tucumaTucumaMod2)

ucuubaAndiroba$ponto <- interaction(
  ucuubaAndiroba$Ucuuba,
  ucuubaAndiroba$Andiroba,
  ucuubaAndiroba$Brijo10
)

modScheffe <- lm(
  PDI ~ Murumuru + Acai + Brijo10 +
    Murumuru:Acai + Murumuru:Brijo10 + Acai:Brijo10 - 1,
  data = murumuruAcai
)

modSat <- lm(
  PDI ~ ponto,
  data = ucuubaAndiroba
)


#Lack of Fit test
anova(ucuubaAndirobaMod1, modSat)

#View(murumuruBuriti)
```

### Results

The **Linear Model** was chosen to fit the data.

## Nano Ucuuba-Buriti

```{r}
#| warning: false
#| message: false

ucuubaBuriti <- read.csv("C:/Users/luisr/Documentos/Documents/Documentos - Copia/NanoIA/planilhas/NanoUcuubaBuriti.csv")

head(ucuubaBuriti)

#ucuubaAndiroba$Andiroba <- ucuubaAndiroba$Pracaxi

ucuubaBuritiVar <- c("Ucuuba", "Buriti", "Brijo10")

ucuubaBuritiMod1 <- MixModel(ucuubaBuriti, "PDI", ucuubaBuritiVar, 1)

summary(ucuubaBuritiMod1)

anova(ucuubaBuritiMod1)

ucuubaBuritiMod2 <- MixModel(ucuubaBuriti, "PDI", ucuubaBuritiVar, 2)

summary(ucuubaBuritiMod2)

anova(ucuubaBuritiMod2)

#tucumaUrucumMod3 <- MixModel(tucumaUrucum, "PDI", tucumaUrucumVar, 4)

#summary(cupuacuPequiMod3)

#anova(cupuacuPequiMod3)

anova(ucuubaBuritiMod1, ucuubaBuritiMod2)

#anova(cupuacuPequiMod1, cupuacuPequiMod3)

par(mfrow = c(2, 2))

plot(ucuubaBuritiMod2)

ucuubaBuriti$ponto <- interaction(
  ucuubaBuriti$Ucuuba,
  ucuubaBuriti$Buriti,
  ucuubaBuriti$Brijo10
)

#modScheffe <- lm(
#  PDI ~ Murumuru + Acai + Brijo10 +
#    Murumuru:Acai + Murumuru:Brijo10 + Acai:Brijo10 - 1,
#  data = murumuruAcai
#)

modSat <- lm(
  PDI ~ ponto,
  data = ucuubaBuriti
)


#Lack of Fit test
anova(ucuubaBuritiMod2, modSat)

#View(murumuruBuriti)
```

### Results

The **Quadratic Model** was chosen to fit the data.
